There once was a grid at ol' Carkeek

### 2016-08-13 Manuscript/v0.0/carkeek_grid.tex
fix textypos and render pdf

From each DNA sample, we amplified an approximately 115 base pair (bp) region of the mitochondrial gene encoding 16S RNA using a two-step polymerase chain reaction (PCR) protocol described by ODONNELL (CITE ODONNELL2016).
In the first step, primers were identical in every reaction (forward: AGTTACYYTAGGGATAACAGCG; reverse: CCGGTCTGAACTCAGATCAYGT); primers in the second step included these same sequences but with 3 variable nucleotides ('NNN') and an index sequence on the 5PRIME end (see CITE PRIMER TABLE/METADATA).
We used the program OligoTag \cite{Coissac2012} to generate 30 unique 6 nucleotide index sequences differing by a minimum Hamming distance of 3 (\nameref{primer_table}).
Indexed primers were assigned to samples randomly, with the identical index sequence on the forward and reverse primer to avoid errors associated with dual-indexed multiplexing (CITE Schnell2015).
In a UV-sterilized hood, we prepared 25 microliter reactions containing 18.375 microliters ultrapure water, 2.5 microliters 10x buffer, 0.625 microliters deoxynucleotide solution (8 millimolar), 1 microliter each forward and reverse primer (10 micromolar, obtained lyophilized from Integrated DNA Technologies (Coralville, IA, USA)), 0.25 microliters Qiagen HotStar Taq polymerase, and 1.25 microliter genomic eDNA template at 1:100 dilution in ultrapure water.
PCR thermal profiles began with an initialization step (95C; 15 min) followed by cycles (40 and 20 for the first and second reaction, respectively) of denaturation (95C; 15 sec), annealing (61C; 30 sec), and extension (72C; 30 sec).
20 identical PCRs were conducted from each DNA extract using non-indexed primers; these were pooled into 4 groups of 5 in order to ensure ample template for the subsequent PCR with indexed primers. % we probably pooled >1 product from PCR1, but figuring this out from our lab notebook would be time consuming and perhaps impossible.
In order to isolate the fragment of interest from primer dimer and other spurious fragments generated in the first PCR, we used the AxyPrep Mag FragmentSelect-I kit with solid-phase reversible immobilization (SPRI) paramagnetic beads at 2.5x the volume of PCR product (Axygen BioSciences, Corning, NY, USA).
A 1:5 dilution in ultrapure water of the product was used as template for the second reaction.
PCR products of the second reaction were purified using the Qiagen MinElute PCR Purification Kit (Qiagen, Hilden, Germany).
Ultrapure water was used in place of template DNA and run along with each batch of PCRs to serve as a negative control for PCR; none of these produced visible bands on an agarose gel.
In total, four separate replicates from each of 31 DNA samples were carried through the two-step PCR process for a total of 124 sequenced PCR products.
These were combined with additional samples from other projects, totaling 345 samples.

@@

Of the ______________ taxa ?(for which we recovered data | detected by our methods), we were able to confirm only one which lacks any microscopic and potentially dispersive life stage, the Shiner Surfperch (Embiotocidae).
Adults and juveniles of this species are strictly associated with shallow, near-shore habitats.
Like all members of the family, they are internal fertilizers that give birth to live young, meaning there are no dispersive gametes or larvae which may travel far from juveniles and adults and be detected by our sampling methods.
C. aggregata DNA was abundant at sites nearest to shore, (M_____\%+-SD of all sequences,), rare or absent elsewhere (CITE FIG).
We interpret this as evidence that the chaotic spatial distribution of eDNA communities (CITE FIG PAM) results from our primers' affinity for many species which at some point exist as microscopic pelagic gametes or larvae.
Communities of microscopic organisms can be exceptionally variable in space over small scales due to oceanographic features such as eddies.
Such fine-scale variation would be missed by our sampling design; denser sampling in concert with current profiling would ?(_______) the role of ?(ocean movement???) on eDNA dispersal.

>>>>

From each DNA sample, we amplified an approximately 115 base pair (bp) region of the mitochondrial gene encoding 16S RNA using a two-step polymerase chain reaction (PCR) protocol described by ODONNELL (CITE ODONNELL2016).
In the first step, primers were identical in every reaction (forward: AGTTACYYTAGGGATAACAGCG; reverse: CCGGTCTGAACTCAGATCAYGT); primers in the second step included these same sequences but with 3 variable nucleotides ('NNN') and an index sequence on the 5PRIME end (see CITE PRIMER TABLE/METADATA).
We used the program OligoTag \cite{Coissac2012} to generate 30 unique 6 nucleotide index sequences differing by a minimum Hamming distance of 3 (CITE primer\_table). %\nameref{primer_table}).
Indexed primers were assigned to samples randomly, with the identical index sequence on the forward and reverse primer to avoid errors associated with dual-indexed multiplexing (CITE Schnell2015).
In a UV-sterilized hood, we prepared 25 microliter reactions containing 18.375 microliters ultrapure water, 2.5 microliters 10x buffer, 0.625 microliters deoxynucleotide solution (8 millimolar), 1 microliter each forward and reverse primer (10 micromolar, obtained lyophilized from Integrated DNA Technologies (Coralville, IA, USA)), 0.25 microliters Qiagen HotStar Taq polymerase, and 1.25 microliter genomic eDNA template at 1:100 dilution in ultrapure water.
PCR thermal profiles began with an initialization step (95C; 15 min) followed by cycles (40 and 20 for the first and second reaction, respectively) of denaturation (95C; 15 sec), annealing (61C; 30 sec), and extension (72C; 30 sec).
20 identical PCRs were conducted from each DNA extract using non-indexed primers; these were pooled into 4 groups of 5 in order to ensure ample template for the subsequent PCR with indexed primers. % we probably pooled >1 product from PCR1, but figuring this out from our lab notebook would be time consuming and perhaps impossible.
In order to isolate the fragment of interest from primer dimer and other spurious fragments generated in the first PCR, we used the AxyPrep Mag FragmentSelect-I kit with solid-phase reversible immobilization (SPRI) paramagnetic beads at 2.5x the volume of PCR product (Axygen BioSciences, Corning, NY, USA).
A 1:5 dilution in ultrapure water of the product was used as template for the second reaction.
PCR products of the second reaction were purified using the Qiagen MinElute PCR Purification Kit (Qiagen, Hilden, Germany).
Ultrapure water was used in place of template DNA and run along with each batch of PCRs to serve as a negative control for PCR; none of these produced visible bands on an agarose gel.
In total, four separate replicates from each of 31 DNA samples were carried through the two-step PCR process for a total of 124 sequenced PCR products.
These were combined with additional samples from other projects, totaling 345 samples.

@@

Of the XXXX taxa ?(for which we recovered data | detected by our methods), we were able to confirm only one which lacks any microscopic and potentially dispersive life stage, the Shiner Surfperch (Embiotocidae).
Adults and juveniles of this species are strictly associated with shallow, near-shore habitats.
Like all members of the family, they are internal fertilizers that give birth to live young, meaning there are no dispersive gametes or larvae which may travel far from juveniles and adults and be detected by our sampling methods.
C. aggregata DNA was abundant at sites nearest to shore, (MXXXXX\%+-SD of all sequences,), rare or absent elsewhere (CITE FIG).
We interpret this as evidence that the chaotic spatial distribution of eDNA communities (CITE FIG PAM) results from our primers' affinity for many species which at some point exist as microscopic pelagic gametes or larvae.
Communities of microscopic organisms can be exceptionally variable in space over small scales due to oceanographic features such as eddies.
Such fine-scale variation would be missed by our sampling design; denser sampling in concert with current profiling would ?(XXXX) the role of ?(ocean movement???) on eDNA dispersal.
